---
title: "Tanzania fertilizer use efficiency, profitability and risks across time"
format: 
    html:
      code-fold: true
      code-tools: true
fig-dpi: 300
fig-width: 8.88
fig-align: center
fig-height: 5

self-contained: true
author: Maxwell Mkondiwa (m.mkondiwa@cgiar.org) and Jordan Chamberlin (j.chamberlin@cgiar.org)
editor: visual
toc: true
toc-location: left
number-sections: true
execute: 
  message: false
  warning: false
  echo: true
---

# Introduction

This notebook provides R code for the paper on estimating the heterogenous treatment effects in Tanzania using non-parametric geoadditive and causal machine learning (ML) models. The analytics focus on addressing the following analysis questions:

# Exploration

```{r}
# package names
packages <- c("gplots", "modelsummary", "grf", "policytree", "ggplot2", "micEcon", "frontier", "dplyr", "tidyr", "knitr", "car", "RColorBrewer", "DT", "rio", "tidyr", "dsfa", "mgcv", "geodata", "sf", "mapview", "dplyr", "terra", "raster", "ggridges", "rio", "BART", "BART", "BayesTree", "bartCause", "plm", "rlearner")

# install packages
# installed_packages <- packages %in% rownames(installed.packages())
# if (any(installed_packages == FALSE)) {
#     install.packages(packages[!installed_packages])
# }

# load packages
invisible(lapply(packages, library, character.only = TRUE))
# install.packages("collapse", repos = "https://fastverse.r-universe.dev")
library(rio)


#tz_lsms_panel <- import("tz_lsms_panel_plot_level_221117.dta")
tz_lsms_panel <- import("tz_lsms_panel.csv")

tz_lsms_panel$yld_[tz_lsms_panel$yld_ > quantile(tz_lsms_panel$yld_, 0.99, na.rm = TRUE)] <- NA
tz_lsms_panel$N_kgha[tz_lsms_panel$N_kgha > quantile(tz_lsms_panel$N_kgha, 0.99, na.rm = TRUE)] <- NA

tz_lsms_panel$plotha[tz_lsms_panel$plotha > quantile(tz_lsms_panel$plotha, 0.99, na.rm = TRUE)] <- NA

tz_lsms_panel$N_kgha_dum <- 0
tz_lsms_panel$N_kgha_dum[tz_lsms_panel$N_kgha > 0] <- 1

tz_lsms_panel$N_kgha_dum <- as.numeric(tz_lsms_panel$N_kgha_dum)

tz_lsms_panel$N_kgha_cond <- tz_lsms_panel$N_kgha
tz_lsms_panel$N_kgha_cond[tz_lsms_panel$N_kgha_dum == 0] <- NA

tz_lsms_panel$N_kgha_cond <- as.numeric(tz_lsms_panel$N_kgha_cond)


tz_lsms_panel=subset(tz_lsms_panel, !(is.na(tz_lsms_panel$yld_)))
tz_lsms_panel=subset(tz_lsms_panel, !(is.na(tz_lsms_panel$N_kgha)))
tz_lsms_panel=subset(tz_lsms_panel, !(is.na(tz_lsms_panel$plotha)))
tz_lsms_panel=subset(tz_lsms_panel, !(is.na(tz_lsms_panel$P_kgha)))
tz_lsms_panel=subset(tz_lsms_panel, !(is.na(tz_lsms_panel$harv_yld_mai)))

tz_lsms_panel$soc_5_15cm=tz_lsms_panel$`soc_5-15cm`
tz_lsms_panel$nitrogen_0_5cm=tz_lsms_panel$`nitrogen_0-5cm`
tz_lsms_panel$sand_0_5cm=tz_lsms_panel$`sand_0-5cm`
tz_lsms_panel$ECN_5_15cm=tz_lsms_panel$`ECN_5-15cm`
tz_lsms_panel$pH_0_5cm=tz_lsms_panel$`pH_0-5cm`


```

# Descriptives table

```{r}
library(modelsummary)

mean_na <- function(x) mean(x, na.rm = TRUE)
sd_na <- function(x) SD(x, na.rm = TRUE)
summary_table_by_year <- datasummary(Heading("N obs") * N + Heading("%") * Percent() + (yld_ + N_kgha_dum  +N_kgha_cond+ N_kgha + P_kgha+plotha + ncrops + hhmem + femhead + headeduc + arain_tot+arain_cv) * (mean_na + sd_na) ~ Factor(year), data = tz_lsms_panel, output = "data.frame")

summary_table_by_year_mean <- datasummary(Heading("N obs") * N + Heading("%") * Percent() + (yld_ + N_kgha_dum  +N_kgha_cond+ N_kgha + P_kgha+plotha + ncrops + hhmem + femhead + headeduc + arain_tot+arain_cv) * (mean_na) ~ Factor(year), data = tz_lsms_panel, output = "data.frame")




library(reactable)
library(htmltools)
library(fontawesome)

htmltools::browsable(
    tagList(
        tags$button(
            tagList(fontawesome::fa("download"), "Download as CSV"),
            onclick = "Reactable.downloadDataCSV('summary_table_by_year', 'summary_table_by_year.csv')"
        ),
        reactable(
            summary_table_by_year,
            searchable = TRUE,
            defaultPageSize = 38,
            elementId = "summary_table_by_year"
        )
    )
)

library(reactable)
library(htmltools)
library(fontawesome)

htmltools::browsable(
    tagList(
        tags$button(
            tagList(fontawesome::fa("download"), "Download as CSV"),
            onclick = "Reactable.downloadDataCSV('summary_table_by_year_mean', 'summary_table_by_year_mean.csv')"
        ),
        reactable(
            summary_table_by_year_mean,
            searchable = TRUE,
            defaultPageSize = 38,
            elementId = "summary_table_by_year_mean"
        )
    )
)

# # By district and by year
# summary_table_by_year_dist <- datasummary(Heading("N obs") * N + Heading("%") * Percent() + (yld_ + N_kgha_g_0  + N_kgha + plotha + P_kgha + ncrops + hhmem + femhead + headeduc + arain_tot) * (mean_na + sd_na) ~ Factor(year*district), data = tz_lsms_panel, output = "data.frame")
# 
# library(reactable)
# library(htmltools)
# library(fontawesome)
# 
# htmltools::browsable(
#     tagList(
#         tags$button(
#             tagList(fontawesome::fa("download"), "Download as CSV"),
#             onclick = "Reactable.downloadDataCSV('summary_table_by_year_dist', 'summary_table_by_year_dist.csv')"
#         ),
#         reactable(
#             summary_table_by_year,
#             searchable = TRUE,
#             defaultPageSize = 38,
#             elementId = "summary_table_by_year_dist"
#         )
#     )
# )


tz_lsms_panel$region=as.factor(tz_lsms_panel$region)
tz_lsms_panel$year=as.factor(tz_lsms_panel$year)

# By district and by year
summary_table_by_year_region <- datasummary(Heading("N obs") * N + Heading("%") * Percent() + (yld_ + N_kgha_dum  +N_kgha_cond+ N_kgha + P_kgha+plotha+ ncrops + hhmem + femhead + headeduc + arain_tot)*(mean_na) ~ year*region, data = tz_lsms_panel, output = "data.frame")

summary_table_by_year_region_mean <- datasummary(Heading("N obs") * N + Heading("%") * Percent() + (yld_ + N_kgha_dum  +N_kgha_cond+ N_kgha + P_kgha+plotha + ncrops + hhmem + femhead + headeduc + arain_tot)*(mean_na) ~ year*region, data = tz_lsms_panel, output = "data.frame")


summary_table_by_year_region_t=t(summary_table_by_year_region)
summary_table_by_year_region_mean_t=t(summary_table_by_year_region_mean)

library(reactable)
library(htmltools)
library(fontawesome)

htmltools::browsable(
    tagList(
        tags$button(
            tagList(fontawesome::fa("download"), "Download as CSV"),
            onclick = "Reactable.downloadDataCSV('summary_table_by_year_region_t', 'summary_table_by_year_region_t.csv')"
        ),
        reactable(
            summary_table_by_year_region_t,
            searchable = TRUE,
            defaultPageSize = 38,
            elementId = "summary_table_by_year_region_t"
        )
    )
)


library(reactable)
library(htmltools)
library(fontawesome)

htmltools::browsable(
    tagList(
        tags$button(
            tagList(fontawesome::fa("download"), "Download as CSV"),
            onclick = "Reactable.downloadDataCSV('summary_table_by_year_region_mean_t', 'summary_table_by_year_region_mean_t.csv')"
        ),
        reactable(
            summary_table_by_year_region_mean_t,
            searchable = TRUE,
            defaultPageSize = 38,
            elementId = "summary_table_by_year_region_mean_t"
        )
    )
)
```

# Graphics

```{r}

library(gplots)
plotmeans(yld_ ~ year, main="Yield heterogeineity across years",xlab="Year", ylab="Maize yield (kg/ha)", data=tz_lsms_panel)

library(gplots)
plotmeans(N_kgha ~ year, main="N per ha heterogeineity across years", data=tz_lsms_panel,xlab="Year", ylab="Average N per ha" )

plotmeans(arain_tot~ year, main="Rainfall across years", data=tz_lsms_panel,xlab="Year", ylab="Average total rainfall (mm)" )

library(easynls)
library(lme4)
library(data.table)
library(ggplot2)

# summary_table_by_year_mean_t=t(summary_table_by_year_mean)
# 
# ggplot(summary_table_by_year_mean) +
#   geom_line(aes(x=N, y=Y, group=year, color=year))
# 
# ggplot(summary_table_by_year_region) +
#   geom_line(aes(x=N, y=Y, group=region, color=region))
```

# Mapping Descriptive Statistics

```{r}







```

# Conventional Production Function Approach

## Linear and non-linear parameter models: e.g., Quadratic

```{r}
library(data.table)
library(ggplot2)
library(easynls)
library(lme4)
library(nlme)

tz_lsms_panel=data.table(tz_lsms_panel)
tz_lsms_panel$year=as.factor(tz_lsms_panel$year)

#tz_lsms_panel_sf_subset=subset(tz_lsms_panel_sf, #select=c("V1","yld_","harv_yld_mai","N_kgha","P_kgha","year","soc_5-15cm","population_density","wc2.1_30s_elev","#sand_0-5cm","nitrogen_0-5cm","ECN_5-15cm","pH_0-5cm"))                                                      

#library(tidyr)
#tz_lsms_panel_sf_subset=tz_lsms_panel_sf_subset %>% drop_na()

# Linear
baseline_ols=lm(yld_~N_kgha+plotha+P_kgha+ncrops+hhmem+femhead+headeduc+arain_tot+region+year,data=tz_lsms_panel)

summary(baseline_ols)

# Lm List by year
baseline_ols_yearList <- lmList(yld_~N_kgha+plotha+P_kgha+ncrops+hhmem+femhead+headeduc+arain_tot|year,tz_lsms_panel)

coef(baseline_ols_yearList)

(ci <- confint(baseline_ols_yearList))
plot(ci,cex=2, cex.lab=3, xlab="Maize yield (kg/ha) response", ylab="Year")


## LINEAR PLATEAU
# library(easynls)
# library(data.table)
# tz_lsms_panel <- data.table(tz_lsms_panel)
# tz_lsms_panel.temp <- tz_lsms_panel[, .("yld_","N_kgha")]
# nls.LP <- nlsfit(tz_lsms_panel.temp, model = 3)
# nls.LP$Model
# lp_model=nls.LP$Parameters
# lp_model
# summary(lp_model)

# QUADRATIC

baseline_quad=lm(yld_~N_kgha+I(N_kgha^2)+P_kgha+I(P_kgha^2)+plotha+ncrops+hhmem+femhead+headeduc+arain_tot+region+year,data=tz_lsms_panel)

summary(baseline_quad)

library(modelsummary)

modelplot(baseline_quad)

# Lm List by year
baseline_quad_lmList <- lmList(yld_~N_kgha+I(N_kgha^2)+P_kgha+I(P_kgha^2)+ncrops+hhmem+femhead+headeduc+arain_tot|year,tz_lsms_panel)

coef(baseline_quad_lmList )

(ci <- confint(baseline_quad_lmList ))
plot(ci,cex=2, cex.lab=3, xlab="Maize yield (kg/ha) response", ylab="Year")

## with interactions
# Lm List by year
baseline_quad_lmList_inter <- lmList(yld_~N_kgha*arain_tot+I(N_kgha^2)*arain_tot+P_kgha+I(P_kgha^2)+ncrops+hhmem+femhead+headeduc|year,tz_lsms_panel)

coef(baseline_quad_lmList_inter )

(ci <- confint(baseline_quad_lmList_inter))
plot(ci,cex=2, cex.lab=3, xlab="Maize yield (kg/ha) response", ylab="Year")


# Soil properties Interacted with soils 
baseline_quad_lmList_inter_soil <- lmList(yld_~N_kgha+I(N_kgha^2)+P_kgha+I(P_kgha^2)+ncrops+hhmem+femhead+headeduc+arain_tot+population_density+wc2.1_30s_elev+sand_0_5cm+nitrogen_0_5cm+soc_5_15cm+ ECN_5_15cm+pH_0_5cm|year,tz_lsms_panel)



coef(baseline_quad_lmList_inter_soil )

(ci <- confint(baseline_quad_lmList_inter_soil))
plot(ci,cex=2, cex.lab=3, xlab="Maize yield (kg/ha) response", ylab="Year")




# Quadratic plateau or linear plateau
library(easynls)
tz_lsms_panel.temp <- tz_lsms_panel[, c("yld_","N_kgha")]
nls.QP <- nlsfit(tz_lsms_panel.temp, model = 4)
nls.QP$Model
mn=nls.QP$Parameters
mn
summary(mn)


```

### Site-year specific Quadratic Only response functions

The site-year specific Quadratic response function can be modeled as At level 1, we have $$ Y_{i} = a_{i} + b_{i}*N + c_{i}*N^2 + \varepsilon_{i}  $$ At level 2, we have $$ 
a_{i} \sim N(a_0, \sigma_{a}^2) \\
b_{i} \sim N(b_0, \sigma_{b}^2) \\
c_{i} \sim N(c_0, \sigma_{c}^2) \\
$$ This model can be estimated with the linear mixed model function `lmer` in R package `lme4`

```{r}
tz_lsms_panel.temp <- tz_lsms_panel[, .(yld_, N_kgha, N2 = N_kgha^2, year)]
lmer.Q <- lmer(yld_ ~ 1 + N_kgha + N2  + (1 | year) + (0 + N_kgha | year) + (0 + N2 | year), data = tz_lsms_panel.temp)
lmer.Q
summary(lmer.Q)
```

Or, althernatively, can be estimated with the non-linear mixed model function `nlme`

```{r}
library(nlme)
nlme.Q <- nlme(yld_ ~ (a + b * N_kgha + c * (N_kgha^2)),
                    data = tz_lsms_panel,
                    fixed = a + b + c ~ 1,
                    random = a + b + c ~ 1,
                    groups = ~ year,
                    start = c(800, 10, -0.001))

nlme.Q
```

### Site-year specific Quadratic-plus-plateau response functions

The site-year specific response function can be modeled using a hierarchical quadratic-plus-plateau model. At level 1, we have $$ Y_{i} = \min(a_{i} + b_{i}*N + c_{i}*N^2, Y_{max}) + \varepsilon_{i}  $$ At level 2, we have $$ 
a_{i} \sim N(a_0, \sigma_{a}^2) \\
b_{i} \sim N(b_0, \sigma_{b}^2) \\
c_{i} \sim N(c_0, \sigma_{c}^2) \\
Y_{max} = a_{i} - b_{i}^2/(4*c_i)
$$

It seems the R function `nlme` would work to estimate this model

```{r}
# Define quadratic-plus-plateau function
mq <- lm(yld_ ~ N_kgha + I(N_kgha^2), data=tz_lsms_panel)
a0 <- coef(mq)[[1]]
b0 <- coef(mq)[[2]]
c0 <- coef(mq)[[3]]
clx0 <- -0.5*b0/c0

# Test nls

# fx.QP <- function(N, a, b, c) {
#   y <- (a + b * N + c * I(N^2)) * (N <= -0.5 * b/c) + (a + I(-b^2/(4 * c))) * (N > -0.5 * b/c)
#   return(y)
# }
# 
# nls.QP <- nls(Y ~ fx.QP(N, a, b, c),
#             start = list(a = a0, b = b0, c = c0), data = dat.Puntel.CC.mean,
#             control = nls.control(maxiter = 6000))


# quadplat <- function(x, a, b, clx) {
#   ifelse(x  < clx, a + b * x   + (-0.5*b/clx) * x   * x, 
#                             a + b * clx + (-0.5*b/clx) * clx * clx)
# }
# 
# nls.QP <- nls(Y ~ quadplat(N, a, b, clx), 
#             start = list(a = a0, b = b0, clx = clx0), data = dat.Puntel.CC.mean, 
#             control = nls.control(maxiter = 6000))

# nlme.QP <- nlme(Y ~ fx.QP(N, a, b, c),
#                     data = dat.Puntel.CC.mean,
#                     fixed = a + b + c ~ 1,
#                     random = a + b + c ~ 1,
#                     groups = ~ year,
#                     start = c(a0, b0, c0))
# 
# nlme.QP

# tz_lsms_panel.nlme <- nlme(yld_ ~ (a + b * N_kgha  + c * I(N_kgha  ^2)) * (N_kgha  <= (-0.5 * b/c)) + (a- I(b^2/(4 * c))) * (N_kgha  > (-0.5 * b/c)),
#                     data = tz_lsms_panel,
#                     fixed = a + b + c ~ 1,
#                     random = a + b + c ~ 1,
#                     groups = ~ year,
#                     start = c(a0, b0, c0))
# tz_lsms_panel.nlme

```

### Bayesian analysis

```{r}
# library(brms)
# 
# tz_lsms_panel$Nsq_kgha=tz_lsms_panel$N_kgha^2
# tz_lsms_panel$Y=tz_lsms_panel$yld_
# 
# f1 <- Y ~ (a+ b*N_kgha+c*(Nsq_kgha))*(N_kgha<=(-0.5*b/c))+(a-(b*b/(4*c)))*(N_kgha>(-0.5*b/c))
# 
# prior_1=c(set_prior("normal(5,1)", nlpar="a"),
# set_prior("normal(0,1)", nlpar="b"),
# set_prior("normal(0,1)", nlpar="c"))
# 
# form=bf(f1,nl=TRUE)+list(a~1|year,b~1|year,c~1|year)
# 
# bayesQP=brm(form,prior=prior_1,data=tz_lsms_panel)
# 
# summary(bayesQP)            
```

## Nonlinear parameter models: geoadditive model

```{r}
set.seed(321)

library(bamlss)

# Linear
f1 <- yld_~N_kgha + plotha + P_kgha + ncrops + hhmem + femhead + headeduc + arain_tot +population_density+wc2.1_30s_elev+sand_0_5cm+nitrogen_0_5cm+soc_5_15cm+ ECN_5_15cm+pH_0_5cm+ region + year

b1 <- bamlss(f1, data = tz_lsms_panel, family = "gaussian", n.iter = 12000, burnin = 2000, thin = 10)


summary(b1)

# Nonlinear
# f2=yld_~s(N_kgha)+s(plotha)+s(P_kgha)+s(ncrops)+s(hhmem)+femhead+headeduc+s(arain_tot)+ti(region)+ti(year)
#
# b2=bamlss(f2,data=tz_lsms_panel,family="gaussian", n.iter=12000, burnin=2000, thin=10)
#
#
# summary(b2)
# plot(b2, ask=FALSE)
# #

f3 <- list(
    yld_ ~ s(N_kgha, by = year) + plotha + P_kgha + ncrops + hhmem + femhead + headeduc + arain_tot +population_density+wc2.1_30s_elev+sand_0_5cm+nitrogen_0_5cm+soc_5_15cm+ ECN_5_15cm+pH_0_5cm,

    sigma ~ s(N_kgha, by = year) + plotha + P_kgha + ncrops + hhmem + femhead + headeduc + arain_tot+population_density+wc2.1_30s_elev+sand_0_5cm+nitrogen_0_5cm+soc_5_15cm+ ECN_5_15cm+pH_0_5cm
)

b3 <- bamlss(f3, data = tz_lsms_panel)


summary(b3)
# plot(b3,ask=FALSE)

plot(b3, cex = 2.5, model = "mu", term = "s(N_kgha,by=year)")
plot(b3, cex = 2.5, model = "sigma", term = "s(N_kgha,by=year")

```

# Causal RF approach

## Binary treatment

```{r}
library(grf)
library(policytree)

tz_lsms_panel_estim_fert <- subset(tz_lsms_panel, select = c("fert1_bin", "yld_", "N_kgha", "plotha", "P_kgha", "ncrops", "hhmem", "femhead", "headeduc", "arain_tot", "population_density", "wc2.1_30s_elev", "sand_0_5cm", "nitrogen_0_5cm", "soc_5_15cm", "ECN_5_15cm", "pH_0_5cm", "region", "year", "lat_modified", "lon_modified"))

tz_lsms_panel_estim_fert$region <- as.numeric(tz_lsms_panel_estim_fert$region)
tz_lsms_panel_estim_fert$year <- as.numeric(tz_lsms_panel_estim_fert$year)

library(tidyr)
tz_lsms_panel_estim_fert <- tz_lsms_panel_estim_fert %>% drop_na()


Y_cf_fert <- as.vector(tz_lsms_panel_estim_fert$yld_)
## Causal random forest -----------------

X_cf_fert <- subset(tz_lsms_panel_estim_fert, select = c("plotha", "ncrops", "hhmem", "femhead", "headeduc", "arain_tot", "population_density", "wc2.1_30s_elev", "sand_0_5cm", "nitrogen_0_5cm", "soc_5_15cm", "ECN_5_15cm", "pH_0_5cm", "region", "year"))

X_firststage_cf_fert <- subset(tz_lsms_panel_estim_fert, select = c("plotha", "ncrops", "hhmem", "femhead", "headeduc", "arain_tot", "population_density", "wc2.1_30s_elev", "sand_0_5cm", "nitrogen_0_5cm", "soc_5_15cm", "ECN_5_15cm", "pH_0_5cm", "region", "year"))


W_cf_fert_binary <- as.vector(tz_lsms_panel_estim_fert$fert1_bin)

# Probability random forest to create weights
W.multi_fert.forest_binary <- regression_forest(X_cf_fert, W_cf_fert_binary,
    equalize.cluster.weights = FALSE,
    seed = 2
)
W.hat.multi.all_fert_binary <- predict(W.multi_fert.forest_binary, estimate.variance = TRUE)$predictions


# Regression forest to get expected responses
Y.multi_fert.forest_binary <- regression_forest(X_cf_fert, Y_cf_fert,
    equalize.cluster.weights = FALSE,
    seed = 2
)

print(Y.multi_fert.forest_binary)

varimp.multi_fert_binary <- variable_importance(Y.multi_fert.forest_binary)
Y.hat.multi.all_fert_binary <- predict(Y.multi_fert.forest_binary, estimate.variance = TRUE)$predictions

# Fit binary causal RF model
multi_fert.forest_binary <- causal_forest(X = X_cf_fert, Y = Y_cf_fert, W = W_cf_fert_binary, W.hat = W.hat.multi.all_fert_binary, Y.hat = Y.hat.multi.all_fert_binary, seed = 2)

varimp.multi_fert_cf_binary <- variable_importance(multi_fert.forest_binary)

# Average treatment effects
multi_fert_ate_binary <- average_treatment_effect(multi_fert.forest_binary, target.sample = "overlap")
multi_fert_ate_binary

multi_fert_binary_calibration <- test_calibration(multi_fert.forest_binary)
multi_fert_binary_calibration

```

## Continuous treatment

```{r}
library(grf)
library(policytree)

tz_lsms_panel_estim_fert <- subset(tz_lsms_panel, select = c("fert1_bin", "yld_", "N_kgha", "plotha", "P_kgha", "ncrops", "hhmem", "femhead", "headeduc", "arain_tot", "population_density", "wc2.1_30s_elev", "sand_0_5cm", "nitrogen_0_5cm", "soc_5_15cm", "ECN_5_15cm", "pH_0_5cm", "region", "year", "lat_modified", "lon_modified"))

tz_lsms_panel_estim_fert$region <- as.numeric(tz_lsms_panel_estim_fert$region)
tz_lsms_panel_estim_fert$year <- as.numeric(tz_lsms_panel_estim_fert$year)

library(tidyr)
tz_lsms_panel_estim_fert <- tz_lsms_panel_estim_fert %>% drop_na()


Y_cf_fert <- as.vector(tz_lsms_panel_estim_fert$yld_)
## Causal random forest -----------------

X_cf_fert <- subset(tz_lsms_panel_estim_fert, select = c("plotha", "ncrops", "hhmem", "femhead", "headeduc", "arain_tot", "population_density", "wc2.1_30s_elev", "sand_0_5cm", "nitrogen_0_5cm", "soc_5_15cm", "ECN_5_15cm", "pH_0_5cm", "region", "year"))

X_firststage_cf_fert <- subset(tz_lsms_panel_estim_fert, select = c("plotha", "ncrops", "hhmem", "femhead", "headeduc", "arain_tot", "population_density", "wc2.1_30s_elev", "sand_0_5cm", "nitrogen_0_5cm", "soc_5_15cm", "ECN_5_15cm", "pH_0_5cm", "region", "year"))


W_cf_fert_continuous <- as.vector(tz_lsms_panel_estim_fert$N_kgha)

# Probability random forest to create weights
W.multi_fert.forest_continuous <- regression_forest(X_cf_fert, W_cf_fert_continuous,
    equalize.cluster.weights = FALSE,
    seed = 2
)
W.hat.multi.all_fert_continuous <- predict(W.multi_fert.forest_continuous, estimate.variance = TRUE)$predictions


# Regression forest to get expected responses
Y.multi_fert.forest_continuous <- regression_forest(X_cf_fert, Y_cf_fert,
    equalize.cluster.weights = FALSE,
    seed = 2
)

print(Y.multi_fert.forest_continuous)

varimp.multi_fert_continuous <- variable_importance(Y.multi_fert.forest_continuous)
Y.hat.multi.all_fert_continuous <- predict(Y.multi_fert.forest_continuous, estimate.variance = TRUE)$predictions

# Fit continuous causal RF model
multi_fert.forest_continuous <- causal_forest(X = X_cf_fert, Y = Y_cf_fert, W = W_cf_fert_continuous, W.hat = W.hat.multi.all_fert_continuous, Y.hat = Y.hat.multi.all_fert_continuous, seed = 2)

varimp.multi_fert_cf_continuous <- variable_importance(multi_fert.forest_continuous)

# Average treatment effects
multi_fert_ate_continuous <- average_treatment_effect(multi_fert.forest_continuous, target.sample = "overlap")
multi_fert_ate_continuous

multi_fert_continuous_calibration <- test_calibration(multi_fert.forest_continuous)
multi_fert_continuous_calibration


library(ggridges)
library(dplyr)
library(ggplot2)

tau.multi_fert.forest_continuous <- predict(multi_fert.forest_continuous, target.sample = "all", estimate.variance = TRUE)

tau.multi_fert.forest_continuous <- as.data.frame(tau.multi_fert.forest_continuous)

tau.multi_fert.forest_X <- data.frame(tz_lsms_panel_estim_fert, tau.multi_fert.forest_continuous)

ggplot(tau.multi_fert.forest_X, aes(x = predictions, y = "", fill = factor(stat(quantile)))) +
    stat_density_ridges(
        geom = "density_ridges_gradient", calc_ecdf = TRUE,
        quantiles = 4, quantile_lines = TRUE
    ) +
    scale_y_discrete(expand = c(0.01, 0)) +
    scale_fill_viridis_d(name = "Quartiles") +
    expand_limits(y = 1) +
    theme_bw(base_size = 16) +
    labs(x = "N use effect", y = "Density")

```

# Exploring heterogeneity in Conditional N Use Efficiencies from CRF

## Causal RF

```{r}
library(ggplot2)
NperHa_CATE_N_plot <- ggplot(subset(tau.multi_fert.forest_X, tau.multi_fert.forest_X$N_kgha > 0), aes(N_kgha, predictions)) +
    geom_smooth(method = "loess", formula = y ~ x, col = "darkblue") +
    labs(x = "Applied N per ha", y = "N use efficiency")

previous_theme <- theme_set(theme_bw(base_size = 16))
NperHa_CATE_N_plot

Plotsize_CATE_N_plot <- ggplot(tau.multi_fert.forest_X, aes(plotha, predictions)) +
    geom_smooth(method = "loess", formula = y ~ x, col = "darkblue") +
    labs(x = "Plot size (ha)", y = "N use efficiency")
previous_theme <- theme_set(theme_bw(base_size = 16))
Plotsize_CATE_N_plot

P_CATE_N_plot <- ggplot(tau.multi_fert.forest_X, aes(P_kgha, predictions)) +
    geom_smooth(method = "loess", formula = y ~ x, col = "darkblue") +
    labs(x = "P", y = "N use efficiency")
previous_theme <- theme_set(theme_bw(base_size = 16))
P_CATE_N_plot

Hhmem_CATE_N_plot <- ggplot(tau.multi_fert.forest_X, aes(hhmem, predictions)) +
    geom_smooth(method = "loess", formula = y ~ x, col = "darkblue") +
    labs(x = "HHMEM", y = "N use efficiency")
previous_theme <- theme_set(theme_bw(base_size = 16))
Hhmem_CATE_N_plot


# By year
NperHa_CATE_N_plot_yr <- ggplot(subset(tau.multi_fert.forest_X, tau.multi_fert.forest_X$N_kgha > 0), aes(N_kgha, predictions)) +
    geom_smooth(method = "loess", formula = y ~ x, col = "darkblue") +
    lims(x = c(0, 100)) +
    labs(x = "Applied N per ha", y = "N use efficiency") +
    theme_bw(base_size = 16) +
    facet_wrap(~year)

NperHa_CATE_N_plot_yr

# By soil organic carbon

library(ggplot2)
soc_CATE_N_plot <- ggplot(subset(tau.multi_fert.forest_X, tau.multi_fert.forest_X$soc_5_15cm > 0), aes(soc_5_15cm, predictions)) +
    geom_smooth(method = "loess", formula = y ~ x, col = "darkblue") +
    labs(x = "Soil organic carbon (%)", y = "N use efficiency")

previous_theme <- theme_set(theme_bw(base_size = 16))
soc_CATE_N_plot

# By soil sand
library(ggplot2)
sand_CATE_N_plot <- ggplot(subset(tau.multi_fert.forest_X, tau.multi_fert.forest_X$sand_0_5cm > 0), aes(sand_0_5cm, predictions)) +
    geom_smooth(method = "loess", formula = y ~ x, col = "darkblue") +
    labs(x = "Sand (%)", y = "N use efficiency")

previous_theme <- theme_set(theme_bw(base_size = 16))
sand_CATE_N_plot

# Electrical conductivity
library(ggplot2)
soil_elec_cond_CATE_N_plot <- ggplot(subset(tau.multi_fert.forest_X, tau.multi_fert.forest_X$ECN_5_15cm > 0), aes(ECN_5_15cm, predictions)) +
    geom_smooth(method = "loess", formula = y ~ x, col = "darkblue") +
    labs(x = "Soil electrical conductivity", y = "N use efficiency")

previous_theme <- theme_set(theme_bw(base_size = 16))
soil_elec_cond_CATE_N_plot

# By density
library(ggplot2)
pop_density_CATE_N_plot <- ggplot(subset(tau.multi_fert.forest_X, tau.multi_fert.forest_X$population_density > 0), aes(population_density, predictions)) +
    geom_smooth(method = "loess", formula = y ~ x, col = "darkblue") +
    labs(x = "Population density", y = "N use efficiency")

previous_theme <- theme_set(theme_bw(base_size = 16))
pop_density_CATE_N_plot

# By elevation

library(ggplot2)
elev_CATE_N_plot <- ggplot(subset(tau.multi_fert.forest_X, tau.multi_fert.forest_X$wc2.1_30s_elev > 0), aes(wc2.1_30s_elev, predictions)) +
    geom_smooth(method = "loess", formula = y ~ x, col = "darkblue") +
    labs(x = "Elevation (masl)", y = "N use efficiency")

previous_theme <- theme_set(theme_bw(base_size = 16))
elev_CATE_N_plot


```

# Are N use efficiencies falling overtime?

## Causal RF

```{r}
Year_CATE_N_plot=ggplot(tau.multi_fert.forest_X,aes(year,predictions))+
  geom_smooth(method="loess",formula=y~x,col="darkblue")+
  labs(x="Year",y="N use efficiency")
previous_theme <- theme_set(theme_bw())
Year_CATE_N_plot


```

# Double machine learning (DML)

```{r}





```

# Mapping the estimates

```{r}
library(ggridges)
library(dplyr)
library(ggplot2)

tau.multi_fert.forest_dummy <- predict(multi_fert.forest_binary, target.sample = "all", estimate.variance = TRUE)

tau.multi_fert.forest_dummy <- as.data.frame(tau.multi_fert.forest_dummy)

tau.multi_fert.forest_X_dummy <- data.frame(tz_lsms_panel_estim_fert, tau.multi_fert.forest_dummy)

ggplot(tau.multi_fert.forest_X_dummy, aes(x = predictions, y = "", fill = factor(stat(quantile)))) +
    stat_density_ridges(
        geom = "density_ridges_gradient", calc_ecdf = TRUE,
        quantiles = 4, quantile_lines = TRUE
    ) +
    scale_y_discrete(expand = c(0.01, 0)) +
    scale_fill_viridis_d(name = "Quartiles") +
    expand_limits(y = 1) +
    theme_bw(base_size = 16) +
    labs(x = "Inorgatic fert use effect", y = "Density")

#
library(sp)
library(sf)
tau.multi_fert.forest_X_dummy_sp <- SpatialPointsDataFrame(cbind(tau.multi_fert.forest_X_dummy$lon_modified, tau.multi_fert.forest_X_dummy$lat_modified), data = tau.multi_fert.forest_X_dummy, proj4string = CRS("+proj=longlat +datum=WGS84"))


library(tmap)
tmap_mode("plot")
Nuse_effect_map <- tm_shape(tau.multi_fert.forest_X_dummy_sp) +
    tm_dots(col = "predictions", title = "Effect of N use on yield (kg/ha)", style = "quantile") +
    tm_layout(legend.outside = TRUE)
Nuse_effect_map
tmap_save(Nuse_effect_map, "figures/Nuse_effect_map.png", width = 600, height = 600, asp = 0)


# N Use efficiency map
library(sp)
library(sf)
tau.multi_fert.forest_X_sp <- SpatialPointsDataFrame(cbind(tau.multi_fert.forest_X$lon_modified, tau.multi_fert.forest_X$lat_modified), data = tau.multi_fert.forest_X, proj4string = CRS("+proj=longlat +datum=WGS84"))


library(tmap)
tmap_mode("plot")
Nuse_efficiency_map <- tm_shape(tau.multi_fert.forest_X_sp) +
    tm_dots(col = "predictions", title = "NUE (kg Maize/Kg N)", style = "quantile") +
    tm_layout(legend.outside = TRUE)
Nuse_efficiency_map
tmap_save(Nuse_efficiency_map, "figures/Nuse_efficiency_map.png", width = 600, height = 600, asp = 0)

# N Use  map
library(sp)
library(sf)

tau.multi_fert.forest_X_sp_small <- subset(tau.multi_fert.forest_X_sp, tau.multi_fert.forest_X_sp$N_kgha > 0)

library(tmap)
tmap_mode("plot")
Nuse_map <- tm_shape(tau.multi_fert.forest_X_sp_small) +
    tm_dots(col = "N_kgha", title = "N applied (Kg/ha)", style = "quantile") +
    tm_layout(legend.outside = TRUE)
Nuse_map
tmap_save(Nuse_map, "figures/Nuse_map.png", width = 600, height = 600, asp = 0)
```

## Map of N Use Efficiencies overtime

```{r}
library(tmap)
tmap_mode("plot")
Nuse_efficiency_map_yr <- tm_shape(tau.multi_fert.forest_X_sp) +
    tm_dots(col = "predictions", title = "NUE (kg Maize/Kg N)", style = "quantile") +
    tm_layout(legend.outside = TRUE) +
    tm_facets("year")
Nuse_efficiency_map_yr
tmap_save(Nuse_efficiency_map_yr, "figures/Nuse_efficiency_map_yr.png", width = 600, height = 600, asp = 0)

```
